---
title: "API_request_project"
author: "Justin Herman"
date: "8/19/2019"
output: html_document
---



## Library import
```{r,message=F,warning=FALSE}
library(httr)
library(jsonlite)
library(lubridate)
library(tidyverse)
library(RMariaDB)
library(DT)
library(knitr)
library(kableExtra)

```

## functions 
```{r}
### Function accesses financial data API and returns a dataframe ready to be inserted into SQL 
### Function takes companies stock tag as an input string   
json_to_df <- function(tag)
    {
    ## Build url
    url <- paste("https://financialmodelingprep.com/api/v3/financials/income-statement/",tag,"?period=quarter",sep="")
    headers = c('Upgrade-Insecure-Requests' = '1')
    params = list(`datatype` = 'json')
    
    ## Make request
    result <- GET(url = url, httr::add_headers(.headers=headers), query = params)
    result<- rawToChar(result$content)
    df <- as.data.frame(fromJSON(result)[2])
    
    ## Fix table column names- remove financials. and special char "."
    colnames(df) <- gsub("financials.|\\.","",colnames(df))
    
    ## Convert Date to datetime/ rename date as reportdate
    df$report_date <- as.Date(df$date, '%Y-%m-%d')
    df <-  df %>%
        select(-date)
    
    ## Build tag column to identify stock ticker
    df$company <- tag
    
    ## Build primary key column convert financial data to numeric
    df$primarykey <- paste(as.character(df$report_date),df$company,sep="-")
    cols.num <- colnames(df)[1:31]
    df[,cols.num] <- sapply(df[cols.num],as.numeric)
    
    ## Data check for NA-
    table(is.na(df))
    
    return(df)
}


### Function initalizes table creation in mysql
### takes stock tag(chracter), tablename(character)
### If table already exists, will tell you to use different name or use update function instead
build_table <- function(tag,tablename){
    
        ## import df from json api call function 
        df <- json_to_df(tag)
    
        ## grab credentials from credential file
        db_credentials<-"C:\\Users\\justin\\Desktop\\xmedia.cnf"
        my_sql_db<-"xmedia"
        
        ## make connection
        my_conn<-dbConnect(RMariaDB::MariaDB(),
                           default.file=db_credentials,
                           group=my_sql_db)

        ## Build table from df
        tryCatch(dbWriteTable(my_conn, value = df, 
                              name = tablename,
                              overwrite =FALSE,
                              row.names = FALSE) ,error= function(e){print("table can not be overwritten and already exists. Please use update table function or change name")})
        
        ## Set primary key to Companytag+Date
        res <- dbSendQuery(my_conn, paste("ALTER TABLE",tablename,"ADD CONSTRAINT websites_pk
                                 PRIMARY KEY (`primarykey`(40)) ;"))
        ## Disconnect
        dbClearResult(res)
        dbDisconnect(my_conn)
}


### loops through list of stock tags and updates SQL DB 
### Will not update db if any of stock data in new queried df already exists in SQL DB
update_table <- function(tags,tablename){
    for (tag in tags){
        
        ## set error checker
        skip_to_next <- FALSE
        
        ## import df from json api call function 
        df <- json_to_df(tag)
        
        ## grab credentials from credential file
        db_credentials<-"C:\\Users\\justin\\Desktop\\xmedia.cnf"
        my_sql_db<-"xmedia"
        
        ## make connection
        my_conn<-dbConnect(RMariaDB::MariaDB(),
                           default.file=db_credentials,
                           group=my_sql_db) 
        
        # insert df into SQL.  Catches primary key conflicts(duplicate data), prints stock wasnt updated, and moves on in loop
        tryCatch(dbWriteTable(my_conn, value = df, 
                          name = tablename, 
                          overwrite= FALSE,   
                          append = TRUE,                         
                          row.names = FALSE),error= function(e){skip_to_next <<- TRUE})
         if(skip_to_next) { print(paste("company",tag, "data already exists in DB"))
                            gc()
                            dbDisconnect(my_conn)
                            next }
        gc()
        dbDisconnect(my_conn)
        print(paste("db was updated correctly with: ",tag ))
    }
}


```



```{r}

top_10_banks <- c('WFC', 'USB', 'PNC', 'BBT', 'STI', 'KEY', 'MTB', 'HBAN', 'ZION', 'CMA', 'FITB',"CFG","RF")
## build a table in sql
build_table('USB',"new_table")
update_table(top_10_banks,"new_table")
```

## start windows function Query

+ Experiment with sql db

```{r}
## practice some query with db
db_credentials<-"C:\\Users\\justin\\Desktop\\xmedia.cnf"
my_sql_db<-"xmedia"
my_conn<-dbConnect(RMariaDB::MariaDB(),
                   default.file=db_credentials,
                   group=my_sql_db) 

## print out description of table
print(dbGetQuery(my_conn, "Describe New_table;"))
```


```{r}
## Check for NA values



res <- dbGetQuery(my_conn, "select * from New_table;")
table(is.na(res))

## look at first 10 rows
res <- dbGetQuery(my_conn, "select * from New_table limit 10;")
datatable(res)
gc()
dbClearResult(res)
dbDisconnect(my_conn)
```


** Build windows functions **

We are looking at 10 of the largest publicly traded banks by cap size.  The data already has a YOY quarterly adjusted growth figure, but I would like to use windows functions to build that figure with our revenue data.  I functionize the code, so that we can run a query that prodcues YOY reswults for any of the numeric columns in the dataset

```{r}

### Uses Lag to build Year over year quarterly growth for desired columns
yty_growth <- function(statistic,table_name){
    my_conn<-dbConnect(RMariaDB::MariaDB(),
                   default.file=db_credentials,
                   group=my_sql_db) 
       query <- paste("WITH last_year_debt  as (SELECT company,report_date,",statistic,", LAG(",statistic, ",4) Over( PARTITION BY company order by report_date)  last_year_",statistic," from ",table_name,") Select company,report_date,",statistic, ",last_year_",statistic,", ROUND((last_year_",statistic,"-",statistic,")/last_year_",statistic,"*-1,3) as growth from last_year_debt;",sep="" )
    #print(query)
    res <- dbSendQuery(my_conn,query)
    my_df <- dbFetch(res)
    print(datatable(my_df))
    dbClearResult(res)
    dbDisconnect(my_conn)
}

yty_growth("Revenue","new_table")
yty_growth("OperatingExpenses","new_table")


### Uses dense rank to rank to first develop a growth percent by any col and then rank those growth rates
### Function inputs are statistic, table name, partition set
### if input partition is set to 1, then function will rank individual company performance
### If partition is not set to 1, function will return ranks of the aggregate banking industry

yty_growth_ranking <- function(statistic,table_name,partition=0){
    ##Build connection
    my_conn<-dbConnect(RMariaDB::MariaDB(),
                   default.file=db_credentials,
                   group=my_sql_db)
    
    ##Build query without partition
    if (partition==0){
    query <- paste("WITH last_year_debt  as (SELECT company,report_date,",statistic,", LAG(",statistic, ",4) Over( PARTITION BY company order by report_date) last_year_",statistic," from ",table_name," ),", statistic,"_table as (Select company,report_date,",statistic,",last_year_",statistic,", ROUND((last_year_",statistic,"-",statistic,")/last_year_",statistic,"*-1,3) as growth from last_year_debt) select *, (case when growth is not null then dense_rank() OVER ( order by growth desc) end) as ranked_growth from ",statistic,"_table;",sep="")
    
    ## Send query request and display result
    res <- dbSendQuery(my_conn,query)
    my_df <- dbFetch(res)
    print(datatable(my_df))
    dbClearResult(res)
    dbDisconnect(my_conn)
    }
    
    ##Build query with partition
    else {
        query <- paste("WITH last_year_debt  as (SELECT company,report_date,",statistic,", LAG(",statistic, ",4) Over( PARTITION BY company order by report_date) last_year_",statistic," from ",table_name," ),", statistic,"_table as (Select company,report_date,",statistic,",last_year_",statistic,", ROUND((last_year_",statistic,"-",statistic,")/last_year_",statistic,"*-1,3) as growth from last_year_debt) select *, (case when growth is not null then dense_rank() OVER ( partition by company order by growth desc) end) as ranked_growth from ",statistic,"_table;",sep="")
    
    ## Send query request and display result
    res <- dbSendQuery(my_conn,query)
    my_df <- dbFetch(res)
    print(datatable(my_df))
    dbClearResult(res)
    dbDisconnect(my_conn)
    }
}


yty_growth_ranking("Revenue","new_table")
yty_growth_ranking("Revenue","new_table",partition=1)


## aggregate data to yearly data and build out running yearly average for last 3 years


```

